### 2025-09-16 YEAR2/AER210_LectureNotes/main.tex
sept 15

\begin{definition}[Definition of Double Integral]
    Let $R$ be a rectangular region in the $xy$-plane given by $R = [a,b] \times [c,d]$. The double integral of a function $f(x,y)$ over the region $R$ is defined as:
    \begin{subequations}
    \begin{equation}
        \iint_R f(x,y) \, dA = \lim_{n \to \infty} \sum_{i=1}^n f(x_i^*,y_i^*) \Delta A_i \quad (\text{Riemann Definition})
    \end{equation}
    where $\Delta A_i$ is the area of the $i$-th subrectangle, and $(x_i^*,y_i^*)$ is a sample point in it. The limit is taken as the maximum diameter of the subrectangles approaches zero.
    \begin{equation}
        \iint_R f(x,y) \, dA = \lim_{n,m \to \infty} \sum_{i=1}^n \sum_{j=1}^m f(x_i^*,y_j^*) \Delta A_{ij} \quad (\text{Grid Formulation})
    \end{equation}
    where $\Delta A_{ij}$ is the area of the $ij$-th subrectangle, and $(x_i^*,y_j^*)$ is a sample point in it. Note that the $\Delta A_{ij}$ may be non-uniform. The limit is taken as the maximum diameter of the subrectangles approaches zero.
    \end{subequations}

@@

    We first consider the first octant, where $x, y, z \ge 0$. Then the total volumn would be eight times the volumn of the first octant. We first express it in polar coordinates:
    $$
        x^2 + y^2 = r^2 \implies z = \sqrt{a^2 - r^2}
    $$
    taking the partial derivatives:
    $$
        \frac{\partial z}{\partial r} = \frac{-r}{\sqrt{a^2 - r^2}}, \quad \frac{\partial z}{\partial \theta} = 0
    $$
    Thus, we can set up the double integral as follows:
    \begin{align*}
        S_{\text{1st octant}} &= \iint_R \sqrt{1 + \left( \frac{\partial z}{\partial r} \right)^2 + \left( \frac{\partial z}{\partial \theta} \right)^2} \, dA \\
        &= \int_0^{\frac{\pi}{2}} \int_0^a \sqrt{1 + \left( \frac{-r}{\sqrt{a^2 - r^2}} \right)^2 + 0^2} \, r \, dr \, d\theta \\
        &= \int_0^{\frac{\pi}{2}} \int_0^a \sqrt{1 + \frac{r^2}{a^2 - r^2}} \, r \, dr \, d\theta \\
        &= \int_0^{\frac{\pi}{2}} \int_0^a \frac{a}{\sqrt{a^2 - r^2}} \, r \, dr \, d\theta \\
        &= \int_0^{\frac{\pi}{2}} \left[ -a \sqrt{a^2 - r^2} \right]_{r=0}^{r=a} d\theta = \int_0^{\frac{\pi}{2}} a^2 \, d\theta = \frac{\pi a^2}{2}
    \end{align*}
    Therefore, the total surface area of the sphere is:
    $$        S = 8 \cdot S_{\text{1st octant}} = 4\pi a^2 $$

\subsection{Triple Integrals in Rectangular Coordinates}
\subsection{Triple Integrals in Cylindrical Coordinates}
\subsection{Triple Integrals in Spherical Coordinates}
\subsection{Change of Variables in Multiple Integrals} \label{sec:change_of_variables}
\chapter{Fluid Mechanics}
\end{document}

>>>>

\begin{definition}[Definition of Double Integral] \label{def:double_integral}
    Let $R$ be a rectangular region in the $xy$-plane given by $R = [a,b] \times [c,d]$. The double integral of a function $f(x,y)$ over the region $R$ is defined as:
    \begin{subequations}
    \begin{equation}
        \iint_R f(x,y) \, dA = \lim_{n \to \infty} \sum_{i=1}^n f(x_i^*,y_i^*) \Delta A_i \quad (\text{Riemann Definition})
    \end{equation}
    where $\Delta A_i$ is the area of the $i$-th subrectangle, and $(x_i^*,y_i^*)$ is a sample point in it. The limit is taken as the maximum diameter of the subrectangles approaches zero.
    \begin{equation}
        \iint_R f(x,y) \, dA = \lim_{n,m \to \infty} \sum_{i=1}^n \sum_{j=1}^m f(x_i^*,y_j^*) \Delta A_{ij} \quad (\text{Grid Formulation})
    \end{equation}
    where $\Delta A_{ij}$ is the area of the $ij$-th subrectangle, and $(x_i^*,y_j^*)$ is a sample point in it. Note that the $\Delta A_{ij}$ may be non-uniform. The limit is taken as the maximum diameter of the subrectangles approaches zero.
    \end{subequations}

@@

    We first consider the first octant, where $x, y, z \ge 0$. Then the total volumn would be eight times the volumn of the first octant. We first express it in polar coordinates:
    $$
        x^2 + y^2 = r^2 \implies z = \sqrt{a^2 - r^2}
    $$
    taking the partial derivatives:
    $$
        \frac{\partial z}{\partial r} = \frac{-r}{\sqrt{a^2 - r^2}}, \quad \frac{\partial z}{\partial \theta} = 0
    $$
    Thus, we can set up the double integral as follows:
    \begin{align*}
        S_{\text{1st octant}} &= \iint_R \sqrt{1 + \left( \frac{\partial z}{\partial r} \right)^2 + \left( \frac{\partial z}{\partial \theta} \right)^2} \, dA \\
        &= \int_0^{\frac{\pi}{2}} \int_0^a \sqrt{1 + \left( \frac{-r}{\sqrt{a^2 - r^2}} \right)^2 + 0^2} \, r \, dr \, d\theta \\
        &= \int_0^{\frac{\pi}{2}} \int_0^a \sqrt{1 + \frac{r^2}{a^2 - r^2}} \, r \, dr \, d\theta \\
        &= \int_0^{\frac{\pi}{2}} \int_0^a \frac{a}{\sqrt{a^2 - r^2}} \, r \, dr \, d\theta \\
        &= \int_0^{\frac{\pi}{2}} \left[ -a \sqrt{a^2 - r^2} \right]_{r=0}^{r=a} d\theta = \int_0^{\frac{\pi}{2}} a^2 \, d\theta = \frac{\pi a^2}{2}
    \end{align*}
    Therefore, the total surface area of the sphere is:
    $$        
        S = 8 \cdot S_{\text{1st octant}} = 4\pi a^2 
    $$   
\end{example}

\begin{example}
    \textit{Let $R$ be the triangular region with vertices at $(0,0,0)$, $(1,0,0)$, and $(0,1,0)$. Find the surface area of the portion of $z = 3x + y^2$ that lies above the region $R$.}

    We first express the function and take the partial derivatives:
    $$
        z = 3x + y^2, \quad \frac{\partial z}{\partial x} = 3, \quad \frac{\partial z}{\partial y} = 2y
    $$
    Thus, we can set up the double integral as follows:
    \begin{align*}
        S &= \iint_R \sqrt{1 + \left( \frac{\partial z}{\partial x} \right)^2 + \left( \frac{\partial z}{\partial y} \right)^2} \, dA \\
        &= \int_0^1 \int_0^{y} \sqrt{1 + 3^2 + (2y)^2} \, dx \, dy \\
        &= \int_0^1 \int_0^{y} \sqrt{10 + 4y^2} \, dx \, dy \\
        &= \int_0^1 \sqrt{10 + 4y^2} \cdot y \, dy \\ 
        \intertext{The extra $y$ is useful for substitution. Let $u = 10 + 4y^2$, then $du = 8y \, dy$ (that's why we integrate w.r.t $x$ first). Thus, we have:}
        S &= \int_{u=10}^{u=14} \sqrt{u} \cdot \frac{du}{8} = \frac{1}{8} \cdot \frac{2}{3} u^{\frac{3}{2}} \Big|_{u=10}^{u=14} \\ 
        &= \frac{1}{12} (14^{\frac{3}{2}} - 10^{\frac{3}{2}}) \approx 1.7
    \end{align*}
\end{example}

\subsection{Triple Integrals}
The idea of a triple integral (in fact n-tuple integral) could be extend from the idea of double integral, similar to 
Definition \ref{def:double_integral}. We can define a triple integral as follows:
\begin{definition}
    Consider a function $f(x,y,z)$ that is continuous a 
    3-D reigion with volumne $V$. We can partition the reigion into $n$ subreigons with volume $\Delta V_i$. Then, we can define the triple integral of $f$ over the reigion $V$ as follows:
    \begin{equation}
        \iiint_V f(x,y,z) \, dV = \lim_{|P| \to 0} \sum_{i=1}^n f(x_i^*, y_i^*, z_i^*) \Delta V_i
    \end{equation}
    where $(x_i^*, y_i^*, z_i^*)$ is a point in the $i$-th subregion, and $|P|$ is the maximum diameter of the subregions. This limit exists and equals to the triple integral if $f$ is continuous on $V$, since the following squeeze theorem holds, similar to the double integral case:
    $$       
     \sum_{i=1}^n m_i \Delta V_i \le \sum_{i=1}^n f(x_i^*, y_i^*, z_i^*) \Delta V_i \le \sum_{i=1}^n M_i \Delta V_i
    $$
    where $m$ and $M$ are the minimum and maximum values of $f$ on $V$, respectively.

    In the retangluar case, we can express $\Delta V_i$ as:
    $$        
    \Delta V_i = \Delta x_i \Delta y_i \Delta z_i
    $$
    Thus we have:
    $$
    dV = dx \, dy \, dz
    $$
\end{definition}

\subsection{Triple Integrals in Rectangular Coordinates}
\begin{definition}[Triple Integral in Rectangular Region]
    Consider a function $f(x,y,z)$ that is continuous over a rectangular box $B$ defined by the inequalities:
    $$
        a \le x \le b, \quad c \le y \le d, \quad r \le z \le s
    $$
    Then, the triple integral of $f$ over the box $B$ can be computed as an iterated integral in any order of integration:
    \begin{equation}
        \iiint_B f(x,y,z) \, dV = \int_a^b \int_c^d \int_r^s f(x,y,z) \, dz \, dy \, dx
    \end{equation}
    or any other permutation of the order of integration.

\begin{example}
    Consider $f(x,y,z)$ over a box $Q = \{(x,y,z) \mid a \le x \le b , c \le y \le d, r \le z \le s\}$. We would form the triple integral as follows:
    $$
        \iiint_Q f(x,y,z) \, dV = \int_a^b \int_c^d \int_r^s f(x,y,z) \, dz \, dy \, dx \\
    $$
    Of course, the order of integration can be change (there are $_3P_3= 6$ ways).
\end{example}

\begin{definition}[Triple Integrals in General Reigion]
    Consider a function $f(x,y,z)$ that is continuous over a general reigion $V$ in the 3-D space. We can describe the reigion $V$ using the following inequalities:
    $$
        a \le x \le b, \quad g_1(x) \le y \le g_2(x), \quad h_1(x,y) \le z \le h_2(x,y)
    $$
    where $g_1(x)$, $g_2(x)$, $h_1(x,y)$, and $h_2(x,y)$ are continuous functions. Then, the triple integral of $f$ over the reigion $V$ can be computed as an iterated integral:
    \begin{equation}
        \iiint_V f(x,y,z) \, dV = \int_a^b \int_{g_1(x)}^{g_2(x)} \int_{h_1(x,y)}^{h_2(x,y)} f(x,y,z) \, dz \, dy \, dx = \iint_R \int_{h_1(x,y)}^{h_2(x,y)} f(x,y,z) \, dz \, dA
    \end{equation}
    where $R = \{(x,y) \mid a \le x \le b, g_1(x) \le y \le g_2(x)\}$ is the projection of the reigion $V$ onto the $xy$-plane.

    The order of integration and their associated bounds can be changed based on the description of the region.
\end{definition}

\begin{example} \label{ex:triple_integral_tetrahedron}
    \textit{Evaluate $\iiint_Q 6xy dV$, where $Q$ is the tetrahedron bounded by the coordinate planes $x=0$, $y=0$, $z=0$ and the plane $2x + y + z = 4$.}

    Let $R$ be the projection of the tetrahedron onto the $xy$-plane. Then, we consider how the tetrahedron change in the $z$ direction over $R$:
    $$
        z = 4 - 2x - y \implies 0 \le z \le 4 - 2x - y
    $$
    So we can set up:
    \begin{align*}
        \iiint_Q 6xy \, dV &= \iint_R \int_0^{4 - 2x - y} 6xy \, dz \, dA \\
        &= \iint_R 6xy(4 - 2x - y) \, dA
    \end{align*}
    Now, we consider how the projection $R$ change in the $y$ direction over $x$:
    $$
        2x + y = 4 \implies 0 \le y \le 4 - 2x, \quad 0 \le x \le 2
    $$
    So we can set up:
    \begin{align*}
        \iiint_Q 6xy \, dV &= \int_0^2 \int_0^{4 - 2x} 6xy(4 - 2x - y) \, dy \, dx \\
        &= \int_0^2 \left[ 12x y^2 - 6x(4 - 2x) y^2 - 2y^3 \right]_{y=0}^{y=4 - 2x} \, dx \\
        &= \int_0^2 (192x - 144x^2 + 24x^3 - 32x^4 + 8x^5) \, dx = \frac{64}{5}
    \end{align*}
\end{example}

\begin{example}
    \textit{Evaluate the integral given in the previous Example \ref{ex:triple_integral_tetrahedron} by intergrating w.r.t. $x$ first.}

    We consider how the tetrahedron change in the $x$ direction over $R$, which is now the projection of the 
    tetrahedron onto the $yz$-plane:
    $$
        2x + y + z = 4 \implies 0 \le x \le \frac{4 - y - z}{2}
    $$
    So we can set up:
    \begin{align*}
        \iiint_Q 6xy \, dV &= \iint_R \int_0^{\frac{4 - y - z}{2}} 6xy \, dx \, dA \\
        &= \frac{3}{4}\iint_R y(4 - y - z)^2 \, dA
    \end{align*}
    Now, we consider how the projection $R$ change in the $y$ direction over $z$:
    $$
        y + z = 4 \implies 0 \le y \le 4 - z, \quad 0 \le z \le 4
    $$
    So we can set up:
    \begin{align*}
        \iiint_Q 6xy \, dV &= \frac{3}{4} \int_0^4 \int_0^{4 - z} y(4 - y - z)^2 \, dy \, dz \\
        &= \frac{64}{5}
    \end{align*}
\end{example}

\begin{example}
    \textit{Find the volume of the solid bounded by the surface $z = 4 - y^2$ and planes given by $x+z = 4$, $x=0$, and $z=0$.}

    First, we select the $z = 4 - y^2$ surface as our base reigion $R$. Then, we consider how the solid change in the $x$ direction over $R$:
    $$
        x + z = 4 \implies 0 \le x \le 4 - z
    $$
    So we can set up:
    \begin{align*}
        V &= \iint_R \int_0^{4 - z} 1 \, dx \, dA \\
        &= \iint_R (4 - z) \, dA
    \end{align*}
    Now, we consider how the projection $R$ change in the $z$ direction over $y$:
    $$        z = 4 - y^2 \implies 0 \le z \le 4 - y^2, \quad -2 \le y \le 2
    $$
    So we can set up:
    \begin{align*}
        V &= \int_{-2}^2 \int_0^{4 - y^2} (4 - z) \, dz \, dy \\
        &= \int_{-2}^2 \left[ 4z - \frac{z^2}{2} \right]_{z=0}^{z=4 - y^2} \, dy \\
        &= \int_{-2}^2 \left( 8 - \frac{y^2}{2} \right) \, dy \\
        &= \frac{128}{5}
    \end{align*}
\end{example}
\subsection{Triple Integrals in Spherical Coordinates}
\subsection{Change of Variables in Multiple Integrals} \label{sec:change_of_variables}
\chapter{Fluid Mechanics}
\end{document}

### 2025-09-16 YEAR2/CHE260_LectureNotes/main.tex
sept 15

\end{shaded}
\section{Systems and Properties in Thermodynamics}
\begin{definition}[Energy]
    You should have learned that energy is the ability to do work.
\end{definition}

@@

These fundamental concepts are in arbitrary units and form the foundation `axioms' in science.

\subsection{Energy}
\paragraph{Energy} An understanding of energy is its ability to \textbf{lift weights}. This is a way we could test if there is energy.
\begin{example}[Potential Energy]
    A 1 kg mass lifted 1 meter has a potential energy of about 10 J. Imagine that it is attached to a balance, with a small mass on the other side, then, it is able to lift that weight.
\end{example}

@@

\paragraph{Steam Cycle} The steam cycle is a thermodynamic cycle that converts heat energy into mechanical work. It consists of four main processes: 
\begin{enumerate}
    \item \textbf{Heating} - The working fluid (steam) is heated in the boiler, converting water into steam and increasing its energy.
    \item \textbf{Expansion} - The high-pressure steam expands in the piston, doing work on the piston and converting thermal energy into mechanical work.
    \item \textbf{Cooling} - The steam is cooled in the condenser, releasing heat to the surroundings and condensing back into water.
    \item \textbf{Compression} - The water is pumped back into the boiler, completing the cycle.
\end{enumerate}

@@

The first law of thermodynamics is simple and easy to understand:
\begin{definition}[First Law of Thermodynamics] \label{def:firstlaw}
    The first law of thermodynamics states that energy cannot be created or destroyed, only transformed from one form to another. In the context of a heat engine, this means that the work output ($W$) is equal to the heat input ($Q_H$) minus the heat rejected ($Q_C$):
    \begin{equation}
        W = Q_H - Q_C
    \end{equation}
\end{definition}
The second law formalizes the concept of the directionality of thermodynamic processes, stating that heat cannot spontaneously flow from a colder body to a hotter body. This law introduces the idea of entropy, a measure of the disorder or randomness in a system, which tends to increase over time in isolated systems.

@@

\subsection{Systems and States}
\begin{definition}[System]
    Any piece of matter or a region of space.
\end{definition}

@@

\subsection{Modeling Ideal Gas}
\begin{definition}[Quasi-Equilibrium Process (Reversible Path)]
    A quasi-equilibrium process is a process that occurs slowly enough that the system remains in a state of equilibrium at all times. In other words, the properties of the system change so (infinitely) slowly that they can be considered constant during the process. In this case, this path is reversible. \\
    (Example) A gas in a cylinder with a movable piston that is compressed or expanded very slowly, allowing the gas to remain in equilibrium throughout the process.
\end{definition}

@@

\section{First Law of Thermodynamics}
\begin{shaded}
\paragraph{Motivation} We want to understand how energy is conserved in a system, and how energy is transferred between a system and its surroundings. This is important for understanding how engines work, how to design efficient systems, and how to analyze thermodynamic processes.
\end{shaded}  
The same definition as before \ref{def:firstlaw}.

@@

\subsection{Work Modes}
\begin{definition}[Boundry Work]
    Force acts on the boundries of a systems and deformes them. \\
    (Example) Expansion and compression of a gas

@@

\paragraph{Curve Fitting in the PV Diagram} In general, we can fit a curve in the $PV$ diagram using the polytropic process equation:
\begin{equation}
    P V^n = \text{constant}
\end{equation}
where $n$ is the polytropic index. Different values of $n$ correspond to different specific types of processes:
\begin{itemize}
    \item $n = 0$: Isobaric process (constant pressure)
    \item $n = 1$: Isothermal process (constant temperature)
    \item $n = \gamma$: Adiabatic process (no heat transfer, where $\gamma = \frac{c_p}{c_v}$ is the heat capacity ratio)
    \item $n \to \infty$: Isochoric process (constant volume)
\end{itemize}
We can compute the work done by the system during a polytropic process as:
\begin{equation}
    w_{12} = -\int_{V_1}^{V_2} P dV = \frac{P_2 V_2 - P_1 V_1}{n - 1} \quad (n \neq 1)
\end{equation}

\begin{definition}[Flow Work]
    Flow work is the work required to push mass into or out of a control volume. It is given by:
    \begin{equation}
        W_{flow} = \frac{PB}{m} = Pv
    \end{equation}
    where $P$ is the pressure, $B$ is the volume flow rate, and $m$ is the mass flow rate. $v$ is the specific volume (volume per unit mass).
    \textbf{Energy flowing a system per unit mass of a fluid} is called \textbf{flow energy} or \textbf{flow work}. It is the energy required to push mass into or out of a control volume. The flow work per unit mass is given by:
    \begin{equation}
       H = U + PV
    \end{equation}
    which is called the \textbf{enthalpy} of the fluid. The enthalpy is an extensive property, so we can define the specific enthalpy as:
    \begin{equation}
        h = u + Pv
    \end{equation}
\end{definition}

\chapter{Heat Transfer}

>>>>

\end{shaded}
% ===================== PART 1: FOUNDATIONS =====================
\section{Foundations and Fundamental Concepts}
\begin{definition}[Energy]
    You should have learned that energy is the ability to do work.
\end{definition}

@@

These fundamental concepts are in arbitrary units and form the foundation `axioms' in science. We first build intuitive physical meaning (ability to lift, cause motion, create change) before formal laws.

\subsection{Mechanical Intuition for Energy Forms}
\paragraph{Energy} An understanding of energy is its ability to \textbf{lift weights}. This is a way we could test if there is energy.
\begin{example}[Potential Energy]
    A 1 kg mass lifted 1 meter has a potential energy of about 10 J. Imagine that it is attached to a balance, with a small mass on the other side, then, it is able to lift that weight.
\end{example}

@@

\paragraph{Steam Cycle (Preview)} The steam cycle is a thermodynamic cycle that converts heat energy into mechanical work. It consists of four main processes (we will formalize cycles later under engines): 
\begin{enumerate}
    \item \textbf{Heating} - The working fluid (steam) is heated in the boiler, converting water into steam and increasing its energy.
    \item \textbf{Expansion} - The high-pressure steam expands in the piston, doing work on the piston and converting thermal energy into mechanical work.
    \item \textbf{Cooling} - The steam is cooled in the condenser, releasing heat to the surroundings and condensing back into water.
    \item \textbf{Compression} - The water is pumped back into the boiler, completing the cycle.
\end{enumerate}

@@

% ===================== PART 2: LAWS OF THERMODYNAMICS (PREVIEW) =====================
\section{Preview of Thermodynamic Laws and Directionality}
The first intuitive principle (later formalized) is conservation of energy in transformations between forms:
\begin{definition}[First Law of Thermodynamics] \label{def:firstlaw}
    The first law of thermodynamics states that energy cannot be created or destroyed, only transformed from one form to another. In the context of a heat engine, this means that the work output ($W$) is equal to the heat input ($Q_H$) minus the heat rejected ($Q_C$):
    \begin{equation}
        W = Q_H - Q_C
    \end{equation}
\end{definition}
The second law formalizes directionality: heat does not spontaneously flow from cold to hot. This introduces \emph{entropy} as a bookkeeping tool for irreversibility. (A rigorous treatment of entropy will follow after we develop state / property concepts.)

@@

% ===================== PART 3: SYSTEM DESCRIPTORS =====================
\section{Systems, States, and Properties}
\begin{definition}[System]
    Any piece of matter or a region of space.
\end{definition}

@@

% ===================== PART 4: IDEAL GAS MODEL =====================
\section{Ideal Gas Model and Molecular Basis}
\begin{definition}[Quasi-Equilibrium Process (Reversible Path)]
    A quasi-equilibrium process is a process that occurs slowly enough that the system remains in a state of equilibrium at all times. In other words, the properties of the system change so (infinitely) slowly that they can be considered constant during the process. In this case, this path is reversible. \\
    (Example) A gas in a cylinder with a movable piston that is compressed or expanded very slowly, allowing the gas to remain in equilibrium throughout the process.
\end{definition}

@@

% ===================== PART 5: FIRST LAW FORMAL DEVELOPMENT =====================
\section{First Law of Thermodynamics (Formal Development)}
\begin{shaded}
\paragraph{Motivation} We want to understand how energy is conserved in a system, and how energy is transferred between a system and its surroundings. This is important for understanding how engines work, how to design efficient systems, and how to analyze thermodynamic processes.
\end{shaded}  
The same definition as before \ref{def:firstlaw}.

@@

% ===================== PART 6: WORK AND ENERGY MODES =====================
\section{Work and Energy Transfer Modes}
\begin{definition}[Boundry Work]
    Force acts on the boundries of a systems and deformes them. \\
    (Example) Expansion and compression of a gas

@@

\paragraph{General Polytropic Representation} In general, we can fit a curve in the $PV$ diagram using the polytropic process equation:
\begin{equation}
    P V^n = \text{constant}
\end{equation}
where $n$ is the polytropic index. Different values of $n$ correspond to different specific types of processes:
\begin{itemize}
    \item $n = 0$: Isobaric process (constant pressure)
    \item $n = 1$: Isothermal process (constant temperature)
    \item $n = \gamma$: Adiabatic process (no heat transfer, where $\gamma = \frac{c_p}{c_v}$ is the heat capacity ratio)
    \item $n \to \infty$: Isochoric process (constant volume)
\end{itemize}
We can compute the work done by the system during a polytropic process as:
\begin{equation}
    w_{12} = -\int_{V_1}^{V_2} P dV = \frac{P_2 V_2 - P_1 V_1}{n - 1} \quad (n \neq 1)
\end{equation}

\begin{definition}[Flow Work]
    Flow work is the work required to push mass into or out of a control volume. It is given by:
    \begin{equation}
        W_{flow} = \frac{PB}{m} = Pv
    \end{equation}
    where $P$ is the pressure, $B$ is the volume flow rate, and $m$ is the mass flow rate. $v$ is the specific volume (volume per unit mass).

    \textbf{Energy flowing a system per unit mass of a fluid} is called \textbf{flow energy} or \textbf{flow work}. It is the energy required to push mass into or out of a control volume. The flow work per unit mass is given by enthalpy, which we will define next.
\end{definition}

\begin{definition}[Enthalpy]
    Enthalpy is a thermodynamic property that represents the total energy of a system, including both its internal energy and the energy associated with its pressure and volume. It is defined as:
    \begin{equation}
        H = U + PV
    \end{equation}
    where $H$ is the enthalpy, $U$ is the internal energy, $P$ is the pressure, and $V$ is the volume. Enthalpy is an extensive property, so we can define the specific enthalpy as:
    \begin{equation}
        h = u + Pv
    \end{equation}
    where $h$ is the specific enthalpy, $u$ is the specific internal energy, and $v$ is the specific volume.

\begin{example}[Constant Pressure Porcess]
    Consider a gas in a cylinder with a movable piston. If we heat the gas at constant pressure, we would haveL
    $$
        P_{\text{gas}} = P_{\text{atm}}
    $$
    The work done by the gas during expansion is given by:
    $$
        w_{12} = -\int_{V_1}^{V_2} P dV = -P (V_2 - V_1)
    $$
    The energy balance for the system can be expressed as:
    $$
        Q_{\text{in}} - W_{\text{out}} = \Delta U_{\text{system}}
    $$
    where $Q_{\text{in}}$ is the heat added to the system, $W_{\text{out}}$ is the work done by the system, and $\Delta U_{\text{system}}$ is the change in internal energy of the system. Rearranging, we have:
    $$
        Q_{\text{in}} = (U_2 - U_1) + (P_2 V_2 - P_1 V_1) = (U_2 + P_2 V_2) - (U_1 + P_1 V_1) = H_2 - H_1
    $$
\end{example}

\begin{theorem}[Enthalpy Change in a Constant Pressure Process]
    For a constant pressure process, the heat added to the system is equal to the change in enthalpy of the system:
    \begin{equation}
        Q_{\text{in}} = H_2 - H_1
    \end{equation}

    For the instantaneous change, we have:
    $$
        \delta Q = du + Pdu
    $$
    from the definition of enthalpy, we have:
    $$
        h = u + Pv \quad \Rightarrow \quad dh = du + PdV + VdP
    $$
    Since the pressure is constant ($dP = 0$), we have:
    $$
        \delta Q = dh
    $$
\end{theorem}

\begin{example}[Constant Volume Process]
    Consider a gas in a cylinder with a fixed piston. If we heat the gas at constant volume, there is no boundary work done by the system, as the volume does not change ($dV = 0$). However, there may be other forms of work done on or by the system, such as electrical work if the gas is heated using an electric heater. The energy balance for the system can be expressed as:
    $$
        Q_{\text{in}} - W_{\text{out}} = \Delta U_{\text{system}}
    $$
    where $Q_{\text{in}}$ is the heat added to the system, $W_{\text{out}}$ is the work done by the system (which is zero in this case), and $\Delta U_{system}$ is the change in internal energy of the system. Rearranging, we have:
    $$
        Q_{\text{in}} = U_2 - U_1
    $$
\end{example}

\begin{theorem}
[Internal Energy Change in a Constant Volume Process]
    For a constant volume process, the heat added to the system is equal to the change in internal energy of the system:
    \begin{equation}
        Q_{\text{in}} = U_2 - U_1
    \end{equation}
    For the instantaneous change, we have:
    $$
        \delta Q = du
    $$  
\end{theorem}

\paragraph{How do we evaluate $\Delta U$ and $\Delta H$?} For an ideal gas, we can use the specific heat capacities at constant volume ($c_v$) and constant pressure ($c_p$) to evaluate the changes in internal energy and enthalpy:
\begin{definition}[Specific Heat]
    The specific heat is the amount of energy required to raise the temperature of a unit mass of a substance by one degree kelvin (or celsius). It is given by:
    \begin{equation}
        c_{\text{average}} = \frac{Q}{\Delta T}
    \end{equation}
    where $Q$ is the heat added to the system and $\Delta T$ is the change in temperature. For the specific heat at each instanentious temperature, we have:
    \begin{equation}
        c(T) = \lim_{\Delta T \to 0} \frac{Q}{\Delta T} = \frac{\delta Q}{dT}
    \end{equation}
    note that $c$ could change with temperature.
\end{definition}

\begin{definition}[Specifc Heat at Constant Volume]
    The specific heat at constant volume ($c_v$) is the amount of energy required to raise the temperature of a unit mass of a substance by one degree kelvin (or celsius) while keeping the volume constant. It is given by:
    \begin{equation}
        c_v(T) = \left(\frac{\partial u}{\partial T}\right)_v 
    \end{equation}
\end{definition}

\begin{definition}[Specific Heat at Constant Pressure]
    The specific heat at constant pressure ($c_p$) is the amount of energy required to raise the temperature of a unit mass of a substance by one degree kelvin (or celsius) while keeping the pressure constant. It is given by:
    \begin{equation}
        c_p(T) = \left(\frac{\partial h}{\partial T}\right)_p 
    \end{equation}
\end{definition}

\paragraph{Ideal Gas} For an ideal gas, the $h$ and $u$ depend only on temperature, so we can write:
\begin{subequations}
\begin{align}
    du &= c_v(T) dT \\
    dh &= c_p(T) dT
\end{align}
\end{subequations}
Also, we have:
$$
    h = u + Pv = u + RT
$$
Differentiating, we have:
$$
    dh = du + R dT
$$
Substituting, we have:
\begin{equation}
    R = c_p - c_v
\end{equation}

\begin{definition}[Specifc Heat Ratio]
    The specific heat ratio ($\gamma$) is the ratio of the specific heat at constant pressure to the specific heat at constant volume:
    \begin{equation}
        \gamma = \frac{c_p}{c_v}
    \end{equation}
    For a monoatomic ideal gas, we have:
    \begin{equation}
        \gamma = \frac{5}{3} \approx 1.67
    \end{equation}
    For a diatomic ideal gas (e.g., nitrogen, oxygen), we have:
    \begin{equation}
        \gamma = \frac{7}{5} = 1.4
    \end{equation}
\end{definition}

\begin{definition}[Change in Internal Energy and Enthalpy]
    The change in internal energy and enthalpy of an ideal gas can be calculated using the specific heat capacities at constant volume ($c_v$) and constant pressure ($c_p$):
    \begin{subequations}
    \begin{align}
        \Delta u &= \int_{T_1}^{T_2} c_v(T) dT \\
        \Delta h &= \int_{T_1}^{T_2} c_p(T) dT
    \end{align}
    \end{subequations}
    where $T_1$ and $T_2$ are the initial and final temperatures, respectively.

    For small temperature changes, since $c_v$ and $c_p$ are relatively constant, we can approximate the changes as:
    \begin{subequations}
        \begin{align}
            \Delta u &\approx c_v (T_2 - T_1) \\
            \Delta h &\approx c_p (T_2 - T_1)
        \end{align}
    \end{subequations}
    To choose the value of $c_v$ and $c_p$, could just use the average value over the temperature range, or use the value at the midpoint temperature.

\begin{example}
    Consider air heated from 300 K to 400 K. For air, we have $c_v = 0.721$ kJ/kg K and $c_p = 1.008$ kJ/kg K. The change in internal energy and enthalpy can be approximated (this is because we assume $c_v$ and $c_p$ are constant over the temperature range, which is a reasonable assumption for small temperature changes) as:

    \begin{align*}
        \Delta u &\approx \int_{T_1}^{T_2} c_v(T) dT = c_v (T_2 - T_1) = 0.721 \times (400 - 300) = 72.1\, \text{kJ/kg} \\
        \Delta h &\approx \int_{T_1}^{T_2} c_p(T) dT = c_p (T_2 - T_1) = 1.008 \times (400 - 300) = 100.8\, \text{kJ/kg} \\
    \end{align*}
\end{example}

\paragraph{Note} Do not associated $c_v$ and $c_p$ with constant volume and constant pressure processes, they are just the specific heat capacities at constant volume and constant pressure, they are properties of the substance that is true for any process. Associate them with $\Delta u$ and $\Delta h$.

\begin{example}
    Consider a piston-cylinder device. If we put a weight on top, the gas is compressed, and the piston moves down. This boundry work process would change enthalpy, despite not changing the temperature. This is becuase:
    $$
        dh = du + PdV + VdP = VdP \quad (du = 0, dV = 0)
    $$
\end{example}
\chapter{Heat Transfer}
